Synthesizing Symmetric Lenses

### 2018-07-11 papers/SymmetricLenses/main.tex
add it

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic.  
%Once that conversion has
%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
This pair of SREs is used to initialize a priority queue ($pq$).
Next, $\PCF{SynthSymLens}$
enters a loop that searches for likely lenses. The loop terminates when the
algorithm believes it is unlikely to find a better lens than the best one it has
found so far (a termination condition defined by $\PCF{\Continue}$). Within each
iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.  The priority of a SRE pair is the
  number of rewrites needed to derive the pair from the originals.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

@@

\subsection{Search Through Types}
The first synthesizer is one that searches through regular expression pairs.
This synthesizer comprises \Expand, \PQ, and \Continue. It proposes regular
expressions that are semantically, but not syntactically, equivalent to the
originals by performing the following rewrites:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{\Probability}
  \end{tabular}
\end{center}
As well as the congruence rules that allow these rewrites to be applied on
subexpressions.

The priority queue yields stochastic regular expressions generated using fewer
rewrites first. Only when there are no more proposed regular expressions derived
from $n$ rewrites will \PCF{\PQ.Pop} propose regular expressions derived from
$n+1$ rewrites.

The procedure \PCF{\Continue} returns false based on the how long the search has
been going, and how hard it expects the next class of problems to be. In
particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c > d + \Log_2(n)
\end{gather*}

When $\Call{Continue}{pq}$ returns $\False$, \Expand stops proposing regular
expression pairs, and instead returns to the user the best lens found thus far.
Note this means that if the algorithm finds a bijective lens, which has zero cost, it
will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs; other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

Just as SREs are REs with probability annotations,
\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
As with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

@@

Just as we analyzed the information content of ordinary regular expressions, we
can analyze the information content of DNF regular expressions. As before, we
use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
of the expected amount of information required to recover a string in
$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
calculate bounds of the expected amount of information required to recover a
string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
$\SDNFLens$.

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
The synthesis procedure comprises three algorithms: one that greedily finds
symmetric DNF lenses (\GreedySynth), one that greedily finds symmetric sequence
lenses (\GreedySeqSynth), and one that finds symmetric atom lenses
(\AtomSynth{}). These three algorithms are hierarchically structured:
\GreedySynth relies on \GreedySeqSynth, \GreedySeqSynth relies on \AtomSynth,
and \AtomSynth rlies on \GreedySynth. The structure of the algorithms mirrors
the structure of symmetric DNF lenses and \SDNFREabbrevs. We describe this
synthesis algorithm using the running example of searching for a lens between
\lstinline{"" | name.name$^*$} and \lstinline{"" | name}.

@@

\GreedySynth loops until there are no more sequence lenses in the priority
queue. Within this loop, a sequence lense is popped from the queue and, if it is
``useful,'' it is included in the final DNF lens. The lens is considered to
be useful when its source (or target) is \textit{not} already the source (or target) of
an already chosen sequence lens. If the examples require that two sequences have a
lens between them, then any sequence lens between them is considered useful. We
update the priorities of the sequence lenses as the algorithm proceeds: if two
sequence lenses have the same source, the second one to be popped gets a higher
cost than it originally had, to account for the information that needs to be stored
for including that source of non-bijectivity.

In our example, \GreedySynth might first pop the sequence pair \lstinline{""} and
\lstinline{""}, because there is a bijective sequence lens between them. As neither
\lstinline{""} is involved in a sequence lens, this lens is considered useful.
Next, the sequence lens between \lstinline{name.name$^*$} and
\lstinline{name} would be popped: while that lens is not bijective: it is still
better than the alternatives. As all sequences are now involved in sequence lenses,
and there are no examples to make other lenses useful, no further sequence
lenses would be added to the lens builder.

@@

\appendix

\ifappendices

@@

\fi
\section{Forgetful Symmetric Lenses}
\begin{theorem*}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem*}
The following lemma is
\begin{proof}
\begin{enumerate}

@@

By the induction hypothesis, $H(SQ_i, p_i) = \mathbb{H}(SQ_i, p_i)$. Using the formula
$$H(P) = H(m(P_1), \ldots m(P_n)) + \sum_{i=1}^n m(P_i)H(P_i)$$
if $P$ is a measure space satisfying $P = P_1 + \ldots + P_n$, then
\begin{align*}
H(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
&= H(p_1, \ldots, p_n) + \sum_{i=1}^n p_i H(SQ_i, p_i)\\
&= \sum_{i=1}^n p_i \mathbb{H}(SQ_i, p_i) + p_i * \log_2 p_i\\
&= \mathbb{H}(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
\end{align*}
\end{enumerate}
\end{proof}
\end{document}

@@

%\end{enumerate}

>>>>

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
level of abstraction. The input regular expressions are first converted into
stochastic regular expressions with \ToStochastic.  
%Once that conversion has
%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
This pair of SREs is used to initialize a priority queue ($pq$). The priority of
a SRE pair is the number of rewrites needed to derive the pair from the
originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely
lenses. The loop terminates when the algorithm believes it is unlikely to find a
better lens than the best one it has found so far (a termination condition
defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
\begin{itemize}
\item pops the next class ($S$, $T$) of lenses to
search off of the priority queue ($\PCF{\PQ.Pop}$),
\item executes $\GreedySynth$ to find a best lens in that class if
one exists ($\LC$),
using the examples $exs$ to filter out potential lenses that do not satisfy
the specification,
\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
to our information theoretic metric, and
\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
  to the priority queue.
\end{itemize}
When the loop terminates, the search returns the globally best lens found
($best$).  Each subroutine of this algorithm will be explained in
further depth in the following subsections.

\emph{Stochastic DNF regular expressions} are critical to the success of our
algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
pseudo-normal form for stochastic regular expressions. Of all the star-semiring
equations required to relate two \SDNFREabbrevs, only those relating to Kleene
star (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs; other axioms, like associativity, commutativity or
distributivity, are unneeded. This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
that have the same type when the Kleene star axioms are unused). This property,
and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
our core algorithms.

@@

\subsection{Searching for $(\Regex,\RegexAlt)$ Candidate Classes}
The first phase of the synthesis algorithm looks for pairs of SREs
$(\Regex,\RegexAlt)$ to drive the \GreedySynth algorithm.  These pairs are
generated using the star unrolling axioms:
\begin{center}
  \begin{tabular}{rcl}
    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\

    $\PRegexStar{\Regex}{\Probability}$
    & \Rewrite
    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{\Probability}
  \end{tabular}
\end{center}
as well as the congruence rules that allow these rewrites to be applied on
subexpressions.
The priority queue yields stochastic regular expressions generated using fewer
rewrites first. Only when there are no more proposed regular expressions derived
from $n$ rewrites will \PCF{\PQ.Pop} propose regular expressions derived from
$n+1$ rewrites.

The procedure \PCF{\Continue} returns false based on the how long the search has
been going, and how hard it expects the next class of problems to be. In
particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$, then:
\begin{gather*}
  \Call{Continue}{pq} := c > d + \Log_2(n)
\end{gather*}

When $\Call{Continue}{pq}$ returns $\False$, the algorithm stops proposing
regular expression pairs, and instead returns to the user the best lens found
thus far. If the algorithm finds a bijective lens, which
has zero cost, it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Just as SREs are REs with probability annotations,
\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
As with SREs and regular
expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
probability annotations.

@@

Just as we analyzed the information content of ordinary regular expressions, we
can analyze the information content of DNF regular expressions. As before, we
use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
of the expected amount of information required to recover a string in
$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
calculate bounds on the expected amount of information required to recover a
string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
$\SDNFLens$.

@@

\subsection{\GreedySynth}
\label{subsec:greedy-synth}
The synthesis procedure comprises three algorithms: one that greedily finds
symmetric DNF lenses (\GreedySynth), one that greedily finds symmetric sequence
lenses (\GreedySeqSynth), and one that finds symmetric atom lenses
(\AtomSynth{}). These three algorithms are hierarchically structured:
\GreedySynth relies on \GreedySeqSynth, \GreedySeqSynth relies on \AtomSynth,
and \AtomSynth rlies on \GreedySynth. The structure of the algorithms mirrors
the structure of symmetric DNF lenses and \SDNFREabbrevs.

@@

\GreedySynth loops until there are no more sequence lenses in the priority
queue. Within this loop, a sequence lens is popped from the queue and, if it is
``useful,'' it is included in the final DNF lens. The lens is considered to
be useful when its source (or target) is \textit{not} already the source (or target) of
an already chosen sequence lens. If the examples require that two sequences have a
lens between them, then any sequence lens between them is considered useful. We
update the priorities of the sequence lenses as the algorithm proceeds: if two
sequence lenses have the same source, the second one to be popped gets a higher
cost than it originally had, to account for the information that needs to be stored
for including that source of non-bijectivity.

As an example, consider searching for a lens between \lstinline{"" | name.name$^*$} and \lstinline{"" | name}. \GreedySynth might first pop the
sequence pair \lstinline{""} and \lstinline{""}, because there is a bijective
sequence lens between them. As neither \lstinline{""} is involved in a sequence
lens, this lens is considered useful. Next, the sequence lens between
\lstinline{name.name$^*$} and \lstinline{name} would be popped: while that lens
is not bijective: it is still better than the alternatives. As all sequences are
now involved in sequence lenses, and there are no examples to make other lenses
useful, no further sequence lenses would be added to the lens builder.

@@

\ifappendices
\appendix

@@

\section{Forgetful Symmetric Lenses}
\begin{theorem*}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem*}
The following lemma is
\begin{proof}
\begin{enumerate}

@@

By the induction hypothesis, $H(SQ_i, p_i) = \mathbb{H}(SQ_i, p_i)$. Using the formula
$$H(P) = H(m(P_1), \ldots m(P_n)) + \sum_{i=1}^n m(P_i)H(P_i)$$
if $P$ is a measure space satisfying $P = P_1 + \ldots + P_n$, then
\begin{align*}
H(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
&= H(p_1, \ldots, p_n) + \sum_{i=1}^n p_i H(SQ_i, p_i)\\
&= \sum_{i=1}^n p_i \mathbb{H}(SQ_i, p_i) + p_i * \log_2 p_i\\
&= \mathbb{H}(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
\end{align*}
\end{enumerate}
\end{proof}

\fi
\end{document}

@@

%  \end{enumerate}
